Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-12 Multiresolution_framework/ltx/manuscript.tex
moved more bits to appendix

For the linear state-space model given by Eqs.~\eqref{eq:StateEquation} and \eqref{eq:ReducedObservationEquation}, the bound defined by Eq.~\eqref{eq:Bishopbound} is a quadratic of the form (see \ref{ap:QDerivation} for derivation) 
\begin{equation}\label{eq:MRA-QCompact}
\mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\varsigma\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\varsigma\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
where $\beta$ is constant with respect to $\boldsymbol\theta$. Equation~\eqref{eq:MRA-QCompact} is maximum at
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top,
\end{align}
where
\begin{align}
\boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Gamma_1]_{j,i}\label{eq:epsilon0} \\
\boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_1]^{j,i} \label{eq:epsilon1}\\
\boldsymbol\Upsilon&=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_2]^{j,i},\label{eq:Epsilon}
\end{align} 
where $[\cdot]_{p,q}$ denotes the $\left(p,q\right)$-element of the matrix, and $ [\cdot]^{p,q}$ denotes the $\left(p,q\right)$-block of the block matrix and where 
\begin{align}
\left[ \Gamma_1\right]^{j,i} &=\sum_{k=1}^{n_x}\left[ \boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1}\right]_{jk} \left[ \mathbf U\right]^{k,i},\\
\left[ \Gamma_2\right] ^{j,i}&=\sum_{k,m=1}^{n_x}[\mathbf U^{\top}]^{jk} \left[\boldsymbol\Lambda_x^{-1}\boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1} \right]_{km}[\mathbf U]^{mi}.
\end{align}
Each $ 1 \times n_{\theta}$ block of the $n_x \times n_x n_{\theta}$ block matrix $\mathbf U$ is 
\begin{align}\label{eq:Uij}
\left[ \mathbf U\right] ^{k,i}&=\int_{\boldsymbol \Omega}\left[\boldsymbol\mu(r) \right]_k \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) \mathrm{d}r'\right]_{i:} \mathrm{d}r,
\end{align}
where $[\cdot]_{i:} $ denotes the $i^{th}$ row. The matrices $\boldsymbol\Xi_0$ and $\boldsymbol\Xi_1$ in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon} are calculated using the RTS smoother outputs: $\hat{\mathbf x}_t$, covariance, $\mathbf P_t^b$, and cross-covariance matrix, $\mathbf M_t$ \citep{Gibsona2005}, 
\begin{align}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \label{eq:Xi0} \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).  \label{eq:Xi1}
\end{align} 
The second derivative of the $\mathcal Q$-function is given by
\begin{equation}
\frac{\partial^2\mathcal Q}{\partial\boldsymbol\theta^2}=-2T_s\varsigma\boldsymbol\Upsilon.
\end{equation}
The matrix $\boldsymbol\Upsilon$ is positive definite if $n_x^2\times n_{\theta}$ matrix $\mathrm{vec}(\mathbf U)$ is of rank $n_{\theta}$ (\citep{Dewar2009}, Lemma 3), where $\mathrm{vec}(\cdot)$ denotes the vectorization operator. Under this condition $\boldsymbol\Upsilon$ is  invertible and the second derivative is negative definite representing a maximum of the $\mathcal Q$-function.

@@

$\boldsymbol\upsilon_0$, $\boldsymbol\upsilon_1$ and $\boldsymbol\Upsilon$ are defined in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon}.

By differentiating the $\mathcal Q$-function with respect to $\boldsymbol\theta$ we have
\begin{align}\label{eq:MRA-QDerivative}
\frac{\partial \mathcal Q}{\partial \boldsymbol\theta}&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-T_s\varsigma(\boldsymbol\Upsilon^\top+\boldsymbol\Upsilon)\boldsymbol\theta \nonumber \\
&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-2T_s\varsigma\boldsymbol\Upsilon^\top\boldsymbol\theta.
\end{align}
Equating Eq.~\eqref{eq:MRA-QDerivative} to $\mathbf 0$ vector yields
\begin{align}\label{eq:app-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top.
\end{align}   
\section{Calculation of $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$}\label{ap:Xiderivation}
The expressions for $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$ in terms of smoother outputs can be found in \cite{Shumway2000}, here we present the derivation for completeness.  The expression for  $\boldsymbol\Xi_{0}$ as defined in Eq.~\eqref{eq:MRA-Xi0} of the main text is given by
   \begin{equation}\label{eq:appXi0}
	 \boldsymbol\Xi_0=\mathbf E_{\theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right].
	\end{equation}                                                                                                
	We start by computing the cross-covariance matrix as

>>>>

For the linear state-space model given by Eqs.~\eqref{eq:StateEquation} and \eqref{eq:ReducedObservationEquation}, the bound defined by Eq.~\eqref{eq:Bishopbound} is a quadratic of the form (see \ref{ap:QDerivation} for derivation) 
\begin{equation}\label{eq:MRA-QCompact}
\mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\varsigma\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\varsigma\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
where $\beta$ is constant with respect to $\boldsymbol\theta$. Equation~\eqref{eq:MRA-QCompact} is maximum at
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top,
\end{align}
where
\begin{align}
\boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Gamma_1]_{j,i}\label{eq:epsilon0} \\
\boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_1]^{j,i} \label{eq:epsilon1}\\
\boldsymbol\Upsilon&=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_2]^{j,i},\label{eq:Epsilon}
\end{align} 
where $[\cdot]_{p,q}$ denotes the $\left(p,q\right)$-element of the matrix, and $ [\cdot]^{p,q}$ denotes the $\left(p,q\right)$-block of the block matrix and where 
\begin{align}
\left[ \Gamma_1\right]^{j,i} &=\sum_{k=1}^{n_x}\left[ \boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1}\right]_{jk} \left[ \mathbf U\right]^{k,i},\\
\left[ \Gamma_2\right] ^{j,i}&=\sum_{k,m=1}^{n_x}[\mathbf U^{\top}]^{jk} \left[\boldsymbol\Lambda_x^{-1}\boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1} \right]_{km}[\mathbf U]^{mi}.
\end{align}
Each $ 1 \times n_{\theta}$ block of the $n_x \times n_x n_{\theta}$ block matrix $\mathbf U$ is 
\begin{align}\label{eq:Uij}
\left[ \mathbf U\right] ^{k,i}&=\int_{\boldsymbol \Omega}\left[\boldsymbol\mu(r) \right]_k \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) \mathrm{d}r'\right]_{i:} \mathrm{d}r,
\end{align}
where $[\cdot]_{i:} $ denotes the $i^{th}$ row. The matrices $\boldsymbol\Xi_0$ and $\boldsymbol\Xi_1$ in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon} are calculated using the RTS smoother outputs: $\hat{\mathbf x}_t$, covariance, $\mathbf P_t^b$, and cross-covariance matrix, $\mathbf M_t$ \citep{Gibsona2005}, 
\begin{align}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \label{eq:Xi0} \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).  \label{eq:Xi1}
\end{align}

@@

$\boldsymbol\upsilon_0$, $\boldsymbol\upsilon_1$ and $\boldsymbol\Upsilon$ are defined in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon}. By differentiating the $\mathcal Q$-function with respect to $\boldsymbol\theta$ we have
\begin{align}\label{eq:MRA-QDerivative}
\frac{\partial \mathcal Q}{\partial \boldsymbol\theta}&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-T_s\varsigma(\boldsymbol\Upsilon^\top+\boldsymbol\Upsilon)\boldsymbol\theta \nonumber \\
&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-2T_s\varsigma\boldsymbol\Upsilon^\top\boldsymbol\theta.
Equating Eq.~\eqref{eq:MRA-QDerivative} to $\mathbf 0$ vector yields
\begin{align}\label{eq:app-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top.
The second derivative of the $\mathcal Q$-function is given by
\frac{\partial^2\mathcal Q}{\partial\boldsymbol\theta^2}=-2T_s\varsigma\boldsymbol\Upsilon.
The matrix $\boldsymbol\Upsilon$ is positive definite if $n_x^2\times n_{\theta}$ matrix $\mathrm{vec}(\mathbf U)$ is of rank $n_{\theta}$ (\citep{Dewar2009}, Lemma 3), where $\mathrm{vec}(\cdot)$ denotes the vectorization operator. Under this condition $\boldsymbol\Upsilon$ is  invertible and the second derivative is negative definite representing a maximum of the $\mathcal Q$-function.  
\section{Calculation of $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$}\label{ap:Xiderivation}
The expressions for $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$ in terms of smoother outputs can be found in \cite{Shumway2000}, here we present the derivation for completeness.  The expression for  $\boldsymbol\Xi_{0}$ as defined in Eq.~\eqref{eq:MRA-Xi0} of the main text is given by
   \begin{equation}\label{eq:appXi0}
	 \boldsymbol\Xi_0=\mathbf E_{\theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right].
	We start by computing the cross-covariance matrix as
